Wireless Automobile Detection, License Plate Processing, and Data Availability Network

### 2014-05-01 paper/paper.tex
Changed date to May 1
Submitted Paper

\title{Wireless Automobile Detection, License Plate Processing, and Data Availability Network}
\author{Kevin Emery, Santiago Gonzalez, Brandon Rodriguez, Taylor Sallee\\ \emph{Undergraduates, EECS Department, Colorado School of Mines}}
\date{April 28, 2014}

@@

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself consists of several subsystems, which are described in detail in section 3.

\section{Related Work}
The ACMx group at CSM has designed a system named Smartlots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station is a Raspberry Pi connected to the internet via Ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The Smartlots system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. This system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. Before we started our project, the ACMx team had created a website that would display the status of the CTLM upper and lower lots once the monitoring system had been deployed. \cite{parkingWiki} is updated periodically with new information regarding project progress. The project described in this paper augments and extends the ACMx project while collaborating with Roy Stillwell (ACMx president and coauthor of \cite{stillwell2013}).

K. V. S. Hari \cite{itsindia} researched the use of magnetometers for detecting and classifying automotive vehicles as part of the Centre for Development of Advanced Computing at Thiruvananthapuram, India. Hari conducted this project in the hopes that a similar system could eventually be used in traffic lights to improve the flow of vehicles. Automobiles are detected using a threshold system similar to that used in \cite{stillwell2013}, and are classified using various features in the vehicle's magnetic signature with one or more sensor axes using clustered dipole models.

@@

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting entry and exit of vehicles from a parking lot. This subsystem is designed to be placed on the side of the road next to each parking lot entrance and exit. Automobiles are detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem is based around the commercially available Arduino Fio 16-bit platform which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee radio to the central base station (described in section 3.3). The packet  contains the direction of the passing vehicle, the processor's temperature, the battery voltage, and a portion of the magnetometer data stream. At the time of detection, a hardware interrupt is sent to the image acquisition Raspberry Pi, to notify it that a picture should be taken by its onboard camera. An easily accessible button is also wired onto the breadboard to allow a human to trigger the onboard camera for testing purposes.

@@

\subsubsection{Image Acquisition}
While the automobile detection subsystem is responsible for correctly identifying when a vehicle enters the lot, the image acquisition subsystem is responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate image. The timing difference between when the Raspberry Pi receives the signal from the Arduino Fio and when the image is captured is set to 500 milliseconds so as to allow the car to completely pass the Arduino Fio and enter the capture area. Only one picture is taken of each car so that capture time and power consumption on the Raspberry Pi are limited. This frees up more resources to be allocated towards image processing.

@@

Once the OpenCV software has processed the image, it is sent from the image acquisition Raspberry Pi to the ACMx server via a Wi-Fi connection. As mentioned in section 5.4, we originally intended to send the compressed images to the base station through the same XBee radio that sends the magnetometer sensor's data, but we discovered that attempting to do this opened a can of worms that was too complicated to solve given the time constraints placed upon us.

@@

The central base station is the point of contact between the sensor network and the server. The data gathered by the Arduino Fio sensors are routed to the base station via XBee radios. The base station, which is itself a Raspberry Pi (see Figure \ref{fig:basestation}, forwards the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station, along with all the other Raspberry Pis in our system, runs the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing Python scripts that take care of forwarding the data via HTTP to the server was simple. The more difficult task was writing code to allow the base station Raspberry Pi to interface with the XBee radio, and listen for packets coming from the XBee attached to the Arduino Fio. The Python code that runs on the base station Raspberry Pi is listed in Appendix 2. The base station is also equipped to receive images via the XBee radio, and forward this data to the server. However, as previously mentioned, the time constraints of the project did not allow us to finish the sensor-side code that would need to run on the image acquisition Raspberry Pis to make image transmission via XBee possible. Therefore, when testing, we simply connected the image acquisition Raspberry Pis to a Wi-Fi internet connection, and POSTed directly to the server, using the RESTful web services already in place.

@@

One commercial provider of ALPR software is Q-Free ASA, which is based in the Netherlands. Q-Free is a top search result when searching the web for ALPR software. Their ALPR product, named Intrada, offers a cloud-based solution that allows users to send the image of the license plate to their servers, and get back the plain text representation of the characters. One unique feature Intrada includes that we did not find in any other package surveyed is the ability to identify the state of origin of the license plate. Intrada does this by analyzing minor differences in fonts used across state borders \cite{intrada2014}.

@@

E-470 uses high speed cameras to capture and extract license plate data in order to bill users of their toll road in the Denver Metro area. We reached out to the E-470 Public Highway Authority to inquire more about their deployment of sensors and ALPR software. Representatives informed us that their toll booths used an ALPR system developed by The 3M Company. Even though E-470 is using an ALPR which automatically detects the characters of a license plate, every image must be reviewed by a human operator before the driver is billed \cite{e470}. In our project, we don't want a human operator to have to confirm each plate - we want to trust the ALPR to do its job.

@@

Before we started this project, the ACMx group had already created the publicly-accessible Smartlots webpage. Therefore, we chose to focus on creating the administrator side of the site, and improving the back-end of the entire site to make it more secure and easier to extend in the future. The homepage of the Smartlots site is shown in Figure \ref{fig:homepage}. You can visit this site at \url{http://acmxlabs.org/smartlots/}

@@

\subsubsection{HTTP Interface}
The base station Raspberry Pi needs some way to communicate its data with the web application. The simplest way to meet this need is to provide an HTTP interface on the server. We use HTTP to handle requests from the front end of the web application, so it made sense to also allow the base station to communicate with our server through HTTP. We designed our backend services to be as RESTful \cite{rest} as possible. We used the PHP library Slim \cite{Slim} to create our REST API. For each table in the database, our API includes code to create, read, update, and delete the items, via HTTP POST, GET, PUT, and DELETE requests, respectively. For example, when a car is detected by the magnetometer sensor, the data is received by the base station Raspberry Pi, which then sends an HTTP POST request to http://acmxlabs.org/smartlots/fioData. Upon receiving this request, the server runs the appropriate PHP code to add new entries to the database tables that store the magnetometer and car count data. Furthermore, when the "View Licenses" web page on the administrator website needs to display a table with all the license plate data, it sends an HTTP GET request to http://acmxlabs.org/smartlots/licenses.

@@

Another feature of the administrator webpage is the page that allows administrators to view sensor data. This page maintains a table showing each sensor, its current and last count, its battery level, and the lot where it is deployed. Although this is a simple page, it provides useful information to the system administrators. Future additions to this page will include a way to add and remove sensors from the database.

@@

The original goal in image processing was to crop the image down from a large image to a small, manageable image of just the license plate. This software does exist, and is used in many applications, such as \cite{du2013} and \cite{e470}. However, this software is very hard to implement in practice. Most applications use contour mapping to detect the license plate, but this relies on having an image where the license plate is one of the prominent features in the image. When contour mapping methods were applied to the images in this project, most commonly the contours detected pebbles on the ground in the concrete or lines in the parking lot instead of license plates. While successful in other applications, it did not prove to be a good method in this instance.

@@

The sensor nodes are fairly finicky right now, and do not always work as expected. Because both the Raspberry Pi and the Arduino Fio share a common power source, the Pis sometimes shut down unexpectedly. The magnetometers do not always pick up metal objects passing over them, and therefore we do not always get data when we should. Because we did not have time to implement sending the images through the XBees, our sensor nodes always have to be in range of a Wi-Fi connection, or plugged into an Ethernet port. This would be unrealistic for a real-world deployment. All in all, our sensor nodes need some work before they will be ready to deploy. The rest of the system above the sensor nodes works consistently, and we can confidently say that once the issues with the sensors are worked out, we will be able to consistently send both magnetometer and image data from the sensors to the base station, and have everything be updated on the website in near-real time.

@@

\subsection{Validating and Adjusting Lot Capacity Data}
We expect that our first deployment of this system may not be 100\% accurate at detecting all vehicles that enter and exit lots. Therefore, we would like to create a system to periodically collect overhead images of the parking lots from a camera mounted on the base station Raspberry Pi. These images would serve as a visual representation of how many cars were actually in the parking lot at the time the image was taken. If this feature is implemented, administrators will be able to periodically compare the overhead images to the collected data from the sensor network to make sure the numbers being computed by the system accurately represent the car counts of the lots. A page would be added to the web interface to allow administrators to pull up the image for any lot with an overhead camera and see the numbers for that lot right next to the image. If any discrepancies are found, the administrator would be able to overwrite the count in the database to reflect reality.

Such an interface would be useful in helping to calibrate the system. For example, if we find that some sensors consistently report less cars in the lot than the image shows, these sensors can be fine-tuned to make sure fewer cars slip through unnoticed. Another improvement to work in this area could include an automated version of this reconciliation process, where a computer program could compare the overhead images to the sensed data and make any necessary corrections to the car count in the database.

@@

\subsection{Xbee Image Transmission}
Currently, our system relies on a Wi-Fi network for transmitting license plate images from the Acquisition Raspberry Pi. Future systems would benefit from having such images funneled and transmitted through the Xbee radio, because transmission range would be significantly improved, and eliminates the need for the sensor node to be in range of a Wi-Fi network in order to function. We envision that this would be implemented by sending a bitstream of the image from the Raspberry Pi to the Arduino Fio through a serial link that would then be packetized by the Arduino Fio, and forwarded to the base station via the Xbee radio. However, such a feature would have to be carefully thought out so as to not obstruct the Arduino Fio's automobile detection operations and not congest the Xbee data link. One concern with transmitting images over the XBee comes from the fact that we did not implement any collision avoidance algorithms, and transmitting the images over the XBees would greatly increase the chance of collisions when multiple sensor nodes are sending data to the base station at the same time.

@@

Image acquisition is done by Raspberry Pis stationed at parking lot entrances. These images are sent to a central base station Raspberry Pi, which then uploads the data to the server via our RESTful API. Once on the server, additional image processing is done to retrieve the characters composing the license plate. This data is then stored in a database for use in the web application.

@@

Once deployed, the Smartlots system, augmented with our additional license plate extraction subsystem and improvements to the web application, will offer the Mines campus an accessible way to manage campus parking lots. Faculty and students will be able to use this to find open parking and save time. Future applications could include interfacing with Parking Services to locate illegally parked vehicles or to gather data regarding campus parking needs.

@@

\section*{Appendix 1 --- Intra-team Work Distribution}

This work described in this project was completed by six people. The four authors of this paper contributed the bulk of the work, and two ACMx group members also helped us, mainly by getting us up to speed on the state of the ACMx project when we started, and continuing to be our points of contact to the ACMx group. The work was broken down as follows:

@@

Taylor Sallee wrote all code for the web application, excluding the front-end "Smartlots" part of the site that was already in place before the start of this project. Roy Stilwell from the ACMx group wrote the code for that part of the site. Taylor's contribution was the entire administrator interface, the RESTful API that services the website and the Raspberry Pis when they make POST requests, the user authentication and PHP security added to all sensitive routes created, the database tables for all data that our group was adding (images, licenses, intrada\_licenses, users), and the PHP code for viewing, creating, updating, and deleting data in these tables. He also collaborated with Andrew Koelling to write the base-station Python code that listens for data from the Arduino Fios and Raspberry Pis, and sends that data to the server via HTTP requests. Some of this code was in place before the project, but Taylor and Andrew refactored almost all of it, and added in the parts to handle the image data, and the part that differentiates the Arduino Fio data from the Raspberry Pi (image) data. Taylor also worked with Santiago and Kevin on the sensor Raspberry Pis, making sure that everything was installed correctly, and testing the top-to-bottom functionality of the system (that is, make sure data is getting from the sensors all the way to the web application). Taylor's work is described in sections 3.3, 3.5, 4.3, and 4.5.

@@

Santiago Gonzalez worked primarily with the hardware needed to get this system working. He built the sensor nodes that include both the Arduino Fio and the Raspberry Pi. He deployed the code written by Roy Stilwell to the Arduino Fios, modifying slightly to suit our purposes. He also did most of the work deploying code to the image Acquisition Raspberry Pis, and he was the one that wrote the code that makes the Pis take an image when the magnetometer detects a car. His work is described in section 3.1.

@@

Drew Koelling was a huge help to our group during this project. He met with us several times during our hackathon, even though he did not have as much stake in the success of this project, given that his grade did not depend on what we got done. He primarily helped us understand the base station code, and the existing web interface, and provided Taylor with the information necessary to create a useful REST API. He also helped rewrite some of that base station code. He was overall a great point of contact between our group and the ACMx group, and he was very helpful and good-natured the whole time. We want to explicitly thank him for all the help he gave us in this project.

>>>>

\title{Wireless Automobile Detection, License Plate Processing, and Data Availability Network}
\author{Kevin Emery, Santiago Gonzalez, Brandon Rodriguez, Taylor Sallee\\ \emph{Undergraduates, EECS Department, Colorado School of Mines}}
\date{May 1, 2014}

@@

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that the license plate monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself consists of several subsystems, which are described in detail in section 3.

\section{Related Work}
The ACMx group at CSM has designed a system named Smartlots to track the number of vehicles in a parking lot. Their system is a wireless sensor network consisting of sensor nodes and a central base station. The sensor nodes are based on the Arduino Fio microcontroller platform equipped with triaxial magnetometers and XBee Pro radios. The base station is a Raspberry Pi connected to the internet via Ethernet. The information collected by the sensor nodes is transmitted to the base station and forwarded to a server running a small web application. The Smartlots system is described in \cite{stillwell2013} and \cite{parkingWiki}. \cite{stillwell2013} describes an implementation of a system to detect automobiles entering and exiting a parking lot using the Arduino Fio and the Raspberry Pi. The ACMx system uses the IEEE 802.15.4 communications standard to communicate automobile detection data from the sensor nodes to the base station. \cite{parkingWiki} describes the ongoing status of the ACMx project. Before we started our project, the ACMx team had created a website that would display the status of the CTLM upper and lower lots once the monitoring system had been deployed. \cite{parkingWiki} is updated periodically with new information regarding project progress. The project described in this paper augments and extends the ACMx project while collaborating with Roy Stillwell (ACMx president and coauthor of \cite{stillwell2013}).

K. V. S. Hari \cite{itsindia} researched the use of magnetometers for detecting and classifying automotive vehicles as part of the Centre for Development of Advanced Computing at Thiruvananthapuram, India. Hari conducted the project in the hopes that a similar system could eventually be used in traffic lights to improve the flow of vehicles. Automobiles are detected using a threshold system similar to that used in \cite{stillwell2013}, and are classified using various features in the vehicle's magnetic signature with one or more sensor axes using clustered dipole models.

@@

\subsection{Automobile Detection}
The automobile detection subsystem provides a means for detecting entry and exit of vehicles from a parking lot. This subsystem is designed to be placed on the side of the road next to each parking lot entrance and exit. Automobiles are detected using a magnetometer which perceives the induced change in the local magnetic field as the metallic structure of the vehicle passes by, as described in \cite{stillwell2013}. The automobile detection subsystem is based around the commercially available Arduino Fio 16-bit platform, which utilizes the ubiquitous Atmel ATMEGA328p microcontroller. When an automobile is detected by the Arduino Fio, a packet is sent through the onboard XBee radio to the central base station (described in section 3.3). The packet  contains the direction of the passing vehicle, the processor's temperature, the battery voltage, and a portion of the magnetometer data stream. At the time of detection, a hardware interrupt is sent to the image acquisition Raspberry Pi, to notify it that a picture should be taken by its onboard camera. An easily accessible button is also wired onto the breadboard to allow a human to trigger the onboard camera for testing purposes.

@@

\subsubsection{Image Acquisition}
While the automobile detection subsystem is responsible for correctly identifying when a vehicle enters the lot, the image acquisition subsystem is responsible for determining when to capture a picture of that car in order to optimize the quality of the license plate image. The timing difference between when the Raspberry Pi receives the signal from the Arduino Fio and when the image is captured is set to 500 milliseconds so as to allow the car to completely pass the Arduino Fio and enter the capture area. Only one picture is taken of each car so that capture time and power consumption on the Raspberry Pi are limited, freeing up more resources to be allocated towards image processing.

@@

Once the OpenCV software has processed the image, it is sent from the image acquisition Raspberry Pi to the ACMx server via a Wi-Fi connection. As mentioned in section 5.4, we originally intended to send the compressed images to the base station through the same XBee radio that sends the magnetometer sensor's data, but we discovered that attempting to do so opened a can of worms that was too complicated to solve given the time constraints placed upon us.

@@

The central base station is the point of contact between the sensor network and the server. The data gathered by the Arduino Fio sensors are routed to the base station via XBee radios. The base station, which is itself a Raspberry Pi (see Figure \ref{fig:basestation}, forwards the data to the ACMx server using HTTP requests. Forwarding data is the main function of the base station. The base station, along with all the other Raspberry Pis in our system, runs the Raspbian operating system \cite{raspbian}, a lightweight version of Linux. Raspbian is optimized for the Raspberry Pi, so writing Python scripts that take care of forwarding the data via HTTP to the server was simple. The more difficult task was writing code to allow the base station Raspberry Pi to interface with the XBee radio, and listen for packets coming from the XBee attached to the Arduino Fio. The Python code that runs on the base station Raspberry Pi is listed in Appendix 2. The base station is also equipped to receive images via the XBee radio, and forward it to the server. However, as previously mentioned, the time constraints of the project did not allow us to finish the sensor-side code that would need to run on the image acquisition Raspberry Pis to make image transmission via XBee possible. Therefore, when testing, we simply connected the image acquisition Raspberry Pis to a Wi-Fi internet connection, and POSTed directly to the server, using the RESTful web services already in place.

@@

One commercial provider of ALPR software is Q-Free ASA, which is based in the Netherlands. Q-Free is a top search result when searching the web for ALPR software. Their ALPR product, named Intrada, offers a cloud-based solution that allows users to send the image of the license plate to their servers, and get back the plain text representation of the characters. One unique feature Intrada includes that we did not find in any other package surveyed is the ability to identify the state of origin of the license plate. Intrada identifies the state by analyzing minor differences in fonts used across state borders \cite{intrada2014}.

@@

E-470 uses high speed cameras to capture and extract license plate data in order to bill users of their toll road in the Denver Metro area. We reached out to the E-470 Public Highway Authority to inquire more about their deployment of sensors and ALPR software. Representatives informed us that their toll booths used an ALPR system developed by The 3M Company. Even though E-470 is using an ALPR that automatically detects the characters of a license plate, every image must be reviewed by a human operator before the driver is billed \cite{e470}. In our project, we don't want a human operator to have to confirm each plate - we want to trust the ALPR to do its job.

@@

Before we started this project, the ACMx group had already created the publicly-accessible Smartlots webpage. Therefore, we chose to focus on creating the administrator side of the site, and improving the back-end of the entire site to make it more secure and easier to extend in the future. The homepage of the Smartlots site is shown in Figure \ref{fig:homepage}. You can visit the site at \url{http://acmxlabs.org/smartlots/}

@@

\subsubsection{HTTP Interface}
The base station Raspberry Pi needs some way to communicate its data with the web application. The simplest way to meet this need is to provide an HTTP interface on the server. We use HTTP to handle requests from the front end of the web application, so it made sense to also allow the base station to communicate with our server through HTTP. We designed our backend services to be as RESTful \cite{rest} as possible. We used the PHP library Slim \cite{Slim} to create our REST API. For each table in the database, our API includes code to create, read, update, and delete the items, via HTTP POST, GET, PUT, and DELETE requests, respectively. For example, when a car is detected by the magnetometer sensor, the data is received by the base station Raspberry Pi, which then sends an HTTP POST request to http://acmxlabs.org/smartlots/fioData. Upon receiving this request, the server runs the appropriate PHP code to add new entries to the database tables that store the magnetometer and car count data. Furthermore, when the `View Licenses' web page on the administrator website needs to display a table with all the license plate data, it sends an HTTP GET request to http://acmxlabs.org/smartlots/licenses.

@@

Another feature of the administrator webpage is the page that allows administrators to view sensor data. This page maintains a table showing each sensor, its current and last count, its battery level, and the lot where it is deployed. Although the sensor data page is simple, it provides useful information to the system administrators. Future additions to this page will include a way to add and remove sensors from the database.

@@

The original goal in image processing was to crop the image down from a large image to a small, manageable image of just the license plate. Software to do this kind of thing does exist, and is used in many applications, such as \cite{du2013} and \cite{e470}. However, software like this is very hard to implement in practice. Most applications use contour mapping to detect the license plate, but contour mapping relies on having an image where the license plate is one of the prominent features in the image. When contour mapping methods were applied to the images in this project, most commonly the contours detected pebbles on the ground in the concrete or lines in the parking lot instead of license plates. While successful in other applications, it did not prove to be a good method in this instance.

@@

The sensor nodes are fairly finicky right now, and do not always work as expected. Because both the Raspberry Pi and the Arduino Fio share a common power source, the Pis sometimes shut down unexpectedly. The magnetometers do not always pick up metal objects passing over them, and therefore we do not always get data when we should. Because we did not have time to implement sending the images through the XBees, our sensor nodes always have to be in range of a Wi-Fi connection, or plugged into an Ethernet port. Having to connect the sensor nodes to the internet would be unrealistic for a real-world deployment. All in all, our sensor nodes need some work before they will be ready to deploy. The rest of the system above the sensor nodes works consistently, and we can confidently say that once the issues with the sensors are worked out, we will be able to consistently send both magnetometer and image data from the sensors to the base station, and have everything be updated on the website in near-real time.

@@

\subsection{Validating and Adjusting Lot Capacity Data}
We expect that our first deployment of this system may not be 100\% accurate at detecting all vehicles that enter and exit lots. Therefore, we would like to create a system to periodically collect overhead images of the parking lots from a camera mounted on the base station Raspberry Pi. These images would serve as a visual representation of how many cars were actually in the parking lot at the time the image was taken. If such a feature were implemented, administrators will be able to periodically compare the overhead images to the collected data from the sensor network to make sure the numbers being computed by the system accurately represent the car counts of the lots. A page would be added to the web interface to allow administrators to pull up the image for any lot with an overhead camera and see the numbers for that lot right next to the image. If any discrepancies are found, the administrator would be able to overwrite the count in the database to reflect reality.

Such an interface would be useful in helping to calibrate the system. For example, if we find that some sensors consistently report less cars in the lot than the image shows, these sensors can be fine-tuned to make sure fewer cars slip through unnoticed. Another improvement to work in this area could include an automated version of the lot reconciliation process, where a computer program could compare the overhead images to the sensed data and make any necessary corrections to the car count in the database.

@@

\subsection{Xbee Image Transmission}
Currently, our system relies on a Wi-Fi network for transmitting license plate images from the Acquisition Raspberry Pi. Future systems would benefit from having such images funneled and transmitted through the Xbee radio, because transmission range would be significantly improved, and transmitting the images via XBee eliminates the need for the sensor node to be in range of a Wi-Fi network in order to function. We envision that a system to accomplish XBee image transmission would be implemented by sending a bitstream of the image from the Raspberry Pi to the Arduino Fio through a serial link that would then be packetized by the Arduino Fio, and forwarded to the base station via the Xbee radio. However, such a feature would have to be carefully thought out so as to not obstruct the Arduino Fio's automobile detection operations and not congest the Xbee data link. One concern with transmitting images over the XBee comes from the fact that we did not implement any collision avoidance algorithms, and transmitting the images over the XBees would greatly increase the chance of collisions when multiple sensor nodes are sending data to the base station at the same time.

@@

Image acquisition is done by Raspberry Pis stationed at parking lot entrances. These images are sent to a central base station Raspberry Pi, which then uploads the data to the server via our RESTful API. Once on the server, additional image processing is done to retrieve the characters composing the license plate. The license plate data is then stored in a database for use in the web application.

@@

Once deployed, the Smartlots system, augmented with our additional license plate extraction subsystem and improvements to the web application, will offer the Mines campus an accessible way to manage campus parking lots. Faculty and students will be able to use the Smartlots system to find open parking and save time. Future applications could include interfacing with Parking Services to locate illegally parked vehicles or to gather data regarding campus parking needs.

@@

\section*{Appendix 1 --- Work Distribution}

The work described in this paper was completed by six people. The four authors of this paper contributed the bulk of the work, and two ACMx group members also helped us, mainly by getting us up to speed on the state of the ACMx project when we started, and continuing to be our points of contact to the ACMx group. The work was broken down as follows:

@@

Taylor Sallee wrote all code for the web application, excluding the front-end Smartlots part of the site that was already in place before the start of this project. Roy Stilwell from the ACMx group wrote the code for that part of the site. Taylor's contribution was the entire administrator interface, the RESTful API that services the website and the Raspberry Pis when they make POST requests, the user authentication and PHP security added to all sensitive routes created, the database tables for all data that our group was adding (images, licenses, intrada\_licenses, users), and the PHP code for viewing, creating, updating, and deleting data in these tables. He also collaborated with Andrew Koelling to write the base-station Python code that listens for data from the Arduino Fios and Raspberry Pis, and sends that data to the server via HTTP requests. Some of this code was in place before the project, but Taylor and Andrew refactored almost all of it, and added in the parts to handle the image data, and the part that differentiates the Arduino Fio data from the Raspberry Pi (image) data. Taylor also worked with Santiago and Kevin on the sensor Raspberry Pis, making sure that everything was installed correctly, and testing the top-to-bottom functionality of the system (that is, make sure data is getting from the sensors all the way to the web application). Taylor's work is described in sections 3.3, 3.5, 4.3, and 4.5.

@@

Santiago Gonzalez worked primarily with the hardware needed to get the described system working. He built the sensor nodes that include both the Arduino Fio and the Raspberry Pi. He deployed the code written by Roy Stilwell to the Arduino Fios, modifying slightly to suit our purposes. He also did most of the work deploying code to the image Acquisition Raspberry Pis, and he was the one that wrote the code that makes the Pis take an image when the magnetometer detects a car. His work is described in section 3.1.

@@

Andrew Koelling was a huge help to our group during this project. He met with us several times during our hackathon, even though he did not have as much stake in the success of this project, given that his grade did not depend on what we got done. He primarily helped us understand the base station code, and the existing web interface, and provided Taylor with the information necessary to create a useful REST API. He also helped rewrite some of that base station code. He was overall a great point of contact between our group and the ACMx group, and he was very helpful and good-natured the whole time. We want to explicitly thank him for all the help he gave us in this project.
